Next Steps

### 2019-02-21 hiding/Naive_bayes_V2b.Rmd
cleanup

---
title: "Naive Bayes Workflow Ver 2.0"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments. I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here). This is quoted from Hansen's original tutorial:
>The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

@@

Lines ~30-205 are the baseline version of the workflow using single words. It correctly classifies comments in the top 4 sub-categories  ~79% of the time. While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.

Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.

Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.

To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In the subsequent iterations I will evaluate:

*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
*  Using n-grams rather than single words
*  Making pair-wise comparison versus comparing multiple groups simultaneously
*  Modifying the variables in the NB analysis itself.

@@

##Initial Setup
```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0 of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

```{}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

@@

```{}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

@@

***
START BREAK POINT - Stemming vs. lemmatization
***

@@

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

@@

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels <- comments_rawVERB[1:7005, ]$type
comments_test_labels <- comments_rawVERB[7006:9340,]$type
```

@@

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(comments_train_labels))
prop.table(table(comments_test_labels))
```

@@

***
START BREAK POINT
***
The above code block incorrectly uses sequential roww, not random rows.  I need to find a way to incorporate randomization.

@@

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

@@

###Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier <- naiveBayes(comments_train, comments_train_labels)
comments_test_pred <- predict(comments_classifier, comments_test)
CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table will list how many of the predicted categories are in their correct categories. Adding diagonally and dividing will calculate the percent accuracy.

```{r}
(44+473+1142+206)/23.35
```

To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{r}
predicted_category <- data.frame(comments_test_pred,comments_test_labels)
head(predicted_category)
write.csv(predicted_category, file = "PredictedCategory_Data.csv")
```

***
START BREAK POINT
***
1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
2. Need to add prediction output.
3. Is dropping numerical values down to categorical yes/no variables removing useful info?

***
***
####\

##Visualization
This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
```{r}
wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
ONE <- subset(comments_raw, type == "1_Basic")
TWO <- subset(comments_raw, type == "2_Writing")
THREE <- subset(comments_raw, type == "3_Technical")
FOUR <- subset(comments_raw, type == "4_Logic")

wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

Show the 5 most frequent words in the data:
```{r}
comment_sack <- TermDocumentMatrix(comments_corpus_clean)
comment_pack <- as.matrix(comment_sack)
comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
head(comment_hack, 10)
```

And the 5 most frequent words from each class:
```{r}
wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

##Permutation Tests of Naive Bayes (NB) Method
Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:

*  Compare 2-grams, 3-grams rather than single words
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
*  Modify the variables within the NB analysis itself. 
    +  Laplace constant (Y/N), Cutoff value for words removed from list.
*  Do/Do not stem the terms
    +  This may remove evidence of tense
*  Remove/Leave stopwords
*  Remove/Leave capitalization
*  Remove/Leave numbers
    +  This may remove evidence of technical descriptors
*  Remove/Leave punctuation
    +  This may remove evidence of questions vs statements

>>>>

---
title: "Naive Bayes Workflow Ver 2.0.b"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

@@

##Summary
This is **Version 2.0.b** of the basic workflow for classifying TA comments using a naive Bayes model. It is optimized for text cleanup. It does not randomize the rows selected or make pair-wise assignments.

The protocol is based on a published tutorial [https://rpubs.com/Billyhansen6/318412](located here).

This quote from Hansen's original tutorial nicely explains the challenge of binary classification:
>The objective of our sample project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

@@

***

The main features and permutations to be evaluated are:

*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
*  Using 1- vs. 2- vs. 3-grams
*  Stepwise paired classification versus comparing multiple groups simultaneously

Additional working notes are located at the end of this page.

@@

##Initial Setup
Calling the required libraries.

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0.b of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

```{r}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

@@

```{r}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

@@

***
START BRANCH POINT - Stemming vs. lemmatization
***

@@

***
END BRANCH POINT
***

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

@@

This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_type <- comments_rawVERB[1:7005, ]$type
comments_test_labels_type <- comments_rawVERB[7006:9340,]$type
```

@@

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(comments_train_labels_type))
prop.table(table(comments_test_labels_type))
```

@@

***
START BREAK POINT
***
The above code block incorrectly uses sequential rows, not random rows.  I need to find a way to incorporate randomization.

@@

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

@@

###Train Model, Predict, Evaluate for Type
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
comments_test_pred_type <- predict(comments_classifier_type, comments_test)
CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.

```{r}
print("Overall TYPE accuracy")
(14+45+1346+19)/23.35
```

###Train Model, Predict, Evaluate for LOCUS
This is the same block as above but uses the "Locus" data column.

```{r}
comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.

```{r}
print("Overall LOCUS accuracy")
(62+64+931+3)/23.35
```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{}
predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
head(predicted_category_type)
write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
```

```{}
predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
head(predicted_category_locus)
write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
```

***
START COMMENTS
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0 correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.

Permutations to Test:

*  Compare 1-grams, 2-grams, 3-grams
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Initial approach will be to split between largest group and balance in each category
    +  Ex.: type is 3-technical vs. not 3_technical
*  Modify the variables within the NB analysis itself. 
    +  Laplace constant (Y/N), Cutoff value for words removed from list.
*  Do/Do not stem the terms
    +  This may remove evidence of tense
    +  Better strategy is to lemmatize
*  Remove/Leave stopwords
    +  Look into alternative terms lists to remove
*  Remove/Leave capitalization
*  Remove/Leave numbers
    +  This may remove evidence of technical descriptors
*  Remove/Leave punctuation
    +  This may remove evidence of questions vs statements

Known Issues:

1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
2. Is dropping numerical values down to categorical yes/no variables removing useful info?
3. How to lemmatize instead of stemming?

Keep the code blocks organized so that testing various permutations is mostly on/off.

Create a SEPARATE version for changing larger blocks.

***
END COMMENTS
***

### 2019-02-21 hiding/Naive_bayes_V2r.Rmd
cleanup

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

@@

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{r}
print("Overall structure accuracy")
(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{}
predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{}
predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

@@

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.

Permutations to test with random subsets version (2.0.r):

*  Compare 1-grams, 2-grams, 3-grams
*  Compare outcomes for 3 independent randomized sets of 1-grams

The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix. The second problem is how to pull the training and test data sets out and process them identically.

```{}
set.seed(1234) 		# ensures the same random set gets picked each time.

comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)

comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
```

The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
1. Create a vector of sequential numbers the same length as the number of data frame rows.
2. Randomize the sequential numbers using set.seed to make pattern reproducible.
3. Add the randomized numbers as a new column to the data frame.
4. Re-order the data table to the randomizer column.

This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.

```{}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

```{}
if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

>>>>

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

@@

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(43+473+1131+184)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{r}
print("Overall structure accuracy")
(22+488+292+518+33+10)/23.35  #Need to find a way to calculate this automatically. 
```

@@

```{r}
predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{r}
predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

@@

```{r}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

```{r}
if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

### 2019-02-21 hiding/project.Rmd
cleanup

---
title: "Project Findings"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

>>>>

---
title: "Project Summary"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

### 2019-02-21 index.Rmd
cleanup

---
title: "Analyzing TA Comments on Student Writing in R"
---

@@

## Background
I am interested in understanding how students develop technical writing skills in biology, and more specifically, how we can apply data science methods to improve that process.

My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers. One of our specific aims is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time. To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments TAs make with student performance.

My project for the Faculty Learning Community for R is partly exploratory, in I am asking: what information is present and can be mined from the TA comments dataset? The project also is partly predictive in that I hope to build a text classification workflow that automates the process of categorizing TA comments on student writing.

#### \   
## Prior Work
In Summer 2018, we extracted and tabulated ~11,000 TA comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses collected during a single semester. Using an iteratively generated codebook, all 11,000 comments were classified in two categories:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: did the comment Point Out an error without providing other help, provide simple Copy Correction, provide declarative General or Specific Information only, or provide Holistic coaching that fosters student thinking and long-term improvement?

#### \   
Frequency counts of unambiguously coded TA comments were sorted into a two-axis contingency table that provided two important insights:

* A snapshot of the most common types of comments TAs made, and 
* A relative estimate of where TAs are placing their greatest effort.

#### \   
## Specific Project Goals
TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using existing Unix scripts. However routinely classifying TA comments by hand is impractical; 10-12,000 comments need to be scored EACH SEMESTER. An automated classification method is needed both for speed and consistency.

My specific project goals were/are to create an R-based text analysis workflow that:

1. **Converts TA comments and metadata into a de-identified "tidy" dataframe for R.** This includes converting the master CSV data set to an anonymized data frame with all names and confidential metadata re-keyed to unique random identifiers.
2. **Uses text features to assign TA comments to the categories defined in our existing codebook**. The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater.

@@

>>>>

---
title: "Automated Classification of TA Comments on Student Writing"
---

@@

## Overview
I am interested in understanding how students develop technical writing skills in biology, and how we might improve that process by applying data science methods.

My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers. One specific aim is to see if changing the structure and focus of **instructors' comments** affects how students' writing develops over time. To assess this we must extract instructor comments from student reports, categorize those comments according to their subject and structure, then correlate the types of comments TAs make with student performance.

#### \   
## Prior Work
In Summer 2018, we extracted and tabulated ~11,000 TA comments from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. Using an iteratively generated codebook, the 11,000 TA comments were sepparately classified in two categories:

* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?

#### \   
Counts of coded TA comments were tabulated into a 2-way contingency table that provided us with a direct measure of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.

While classifying TA comments this way is informative, doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.

@@

##Specific Project Goals
My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data. This automated classifier will be used as part of the larger NSF-funded research project to:

*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
*  Monitor how the types and focus of TAs comments on student writing change over time.

#### \   
## Site Map
Insert a text-based site map of the major pages and branch points here.

### 2019-02-21 nextsteps.Rmd
cleanup

---
title: "Background for the Project"
bibliography: library.bib
biblio-style: apalike
---

@@

Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

In 2014 we launched an automated data collection process we call **SAWHET** that systematically collects structured writing and other data from student writers and graduate instructors. Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space. From this empirical basis our practical goal is to develop techniques for teaching technical writing that:

*  Can be implemented at large scale by less experienced teaching assistants;
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
*  Embody other best practices from 40+ years of research literature on "writing in the disciplines."

We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses. In addition we have all comments that a TA made on the individual reports, plus associated metadata:

*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?

We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features. These insights then inform how we teach technical writing and train graduate instructors.

@@

## Context for This Project & Prior Work
Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

In Summer 2018, we began analyzing TAs' comments from student papers. We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses. Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent. In all, the final tabulated dataset included ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

The sorted contingency table provided us with a relative estimate of TA effort.

## The Central Problem of This Project
Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

>>>>

---
title: "Next Steps"
bibliography: library.bib
biblio-style: apalike
---

@@

Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

@@
